#Reproducible research: Course project 

```{r}
library(tidyverse)
library(ggplot2)
```

Load the data 
```{r}
Data <- read.csv("activity.csv", header = TRUE)
head(Data)
```

Calculate the total number of steps per a day and make a histogram 
```{r}
steps_per_day <- Data %>%
      group_by(date) %>%
      summarize(sum_steps = sum(steps, na.rm = TRUE)) 

hist(steps_per_day$sum_steps, 
     main = "Histogram of the total number of steps per a day ", xlab="Steps")
```

Calculate the mean and median of the total step number per a day 
```{r}
pre_NA_mean_stpes_per_day <- mean(steps_per_day$sum_steps)
pre_NA_median_stpes_per_day <- median(steps_per_day$sum_steps)
print(paste("The mean is: ", pre_NA_mean_stpes_per_day))
print(paste("The median is: ", pre_NA_median_stpes_per_day))
```

What is the average daily activity pattern?
Time series plot of 5-mins interval (x-axis) and the average step number 
```{r}
steps_interval <- Data %>%
      group_by(interval) %>%
      summarize(meansteps = mean(steps, na.rm = TRUE)) 

plot(meansteps ~ interval, data = steps_interval, type="l",
     xlab = "5 minute intervals", ylab = "Average step number",
     main = "Steps by time interval")
```

Calculate the interval that contains the most stpes on average 
```{r}
steps_interval$interval[which.max(steps_interval$meansteps)]
print(paste("Interval containing the most steps on average: ",
            steps_interval$interval[which.max(steps_interval$meansteps)]))
```

Calculate and report the total number of missing values in the dataset
```{r}
sum(is.na(Data$steps))
print(paste("The total number of rows with NA is: ",sum(is.na(Data$steps))))
```

Create new data set to fill NA with the mean of 5 mins interval 
```{r}
replacewithmean <- function(x) replace(x, is.na(x), mean(x, na.rm = TRUE))
replace_na_by_mean <- Data%>% 
      group_by(interval) %>% 
      mutate(steps = replacewithmean(steps))
head(replace_na_by_mean)
summary(replace_na_by_mean)
```

Make a histogram and calcuate the mean and median after replace NA
```{r}
steps_per_day_replace_na_by_mean <- replace_na_by_mean %>%
      group_by(date) %>%
      summarize(sum_steps = sum(steps, na.rm = TRUE)) 
hist(steps_per_day_replace_na_by_mean$sum_steps, 
     main = "Histogram of total step numbers per a day ", xlab="Steps")

post_NA_mean_stpes_per_day<-mean(steps_per_day_replace_na_by_mean$sum_steps)
post_NA_median_stpes_per_day <- median(steps_per_day_replace_na_by_mean$sum_steps)

print(paste("The mean after na replacement is: ", post_NA_mean_stpes_per_day))
print(paste("The median after na replacement is: ", post_NA_median_stpes_per_day))
```

Compare before vs after NA replace 
```{r}
print(paste("pre_mean is: ",pre_NA_mean_stpes_per_day))
print(paste("post_mean is:", post_NA_mean_stpes_per_day))

print(paste("pre_median is: ",pre_NA_median_stpes_per_day))
print(paste("post_median is:", post_NA_median_stpes_per_day))
```

Create a new factor variable in the dataset with two levels – “weekday” and “weekend” 
```{r}
replace_na_by_mean$date <- as.Date(replace_na_by_mean$date)
replace_na_by_mean$weekday <- weekdays(replace_na_by_mean$date)
replace_na_by_mean$weekend <- ifelse(replace_na_by_mean$weekday=="Saturday"|replace_na_by_mean$weekday=="Sunday",
                                     "Weekend", "Weekday" )
filtered_weekday <- filter(replace_na_by_mean, weekend == "Weekday")
filtered_weekend <- filter(replace_na_by_mean, weekend == "Weekend")
head(filtered_weekday)
head(filtered_weekend)
```

Make plot
```{r}
steps_interval_weekday <- filtered_weekday %>%
      group_by(interval) %>%
      summarize(meansteps = mean(steps, na.rm = TRUE)) 
steps_interval_weekday$day <- "weekday"

steps_interval_weekend <- filtered_weekend %>%
      group_by(interval) %>%
      summarize(meansteps = mean(steps, na.rm = TRUE)) 
steps_interval_weekend$day <- "weekend"

steps_interval_week_weekend <-rbind(steps_interval_weekday, steps_interval_weekend)
week_weekend <- ggplot (steps_interval_week_weekend, aes (interval, meansteps))+
      geom_line() + facet_grid (day~.) + 
      theme(axis.text = element_text(size = 12),axis.title = element_text(size = 14)) + 
      labs(y = "Step Number") + labs(x = "Interval") + 
      ggtitle("Average step Number")
week_weekend
```

